Safe Transferable Regions

### 2015-11-13 paper/abstract.tex
Updates to abstract

Recent work in high-performance systems written in managed languages (such as Java or C\#) has shown
that garbage-collection can become a significant performance bottleneck in such systems.
A class of these systems, which we refer to as dataflow programs, create huge data-structures with well-defined
lifetimes that open up the possibility for significant performance improvements by using alternative memory
management techniques that can exploit knowledge of these data-structures' lifetimes.

In this paper, we present a memory management scheme that uses user-managed
memory regions that allow programmers to exploit the specific memory behaviours exhibited by dataflow programs.
To protect programmers from the dangers of manual memory management, we
supplement our memory management scheme with a region type system that
statically ensures the memory safety of region-annotated C\# programs,
even in presence of higher-order stateful functions. Further, we
describe a type inference algorithm that infers principal region types
for first-order programs thereby saving programmers the need to write
region annotations on types. Our inference algorithm is also very
effective, albeit incomplete, for higher-order programs.

>>>>

Recent work in high-performance systems written in managed languages (such as Java or C\#) has shown
that garbage-collection can be a significant performance bottleneck in such systems.
A class of these systems, focused on big-data, create huge data-structures with well-defined lifetimes.
In this paper, we present a language and memory management scheme based on user-managed
memory regions (called \emph{transferable regions}) that allow programmers to exploit
knowledge of data-structures' lifetimes to achieve significant performance improvements.

Manual memory management is susceptible to the usual perils of dangling pointers.
A key contribution of this paper is a region type system that
ensures the memory safety of region-annotated C\# programs,
even in presence of higher-order stateful functions. We
also present a type inference algorithm that infers principal region types
for first-order programs, eliminating the need for programmers to write
region annotations on types. Our inference algorithm is also very
effective, albeit incomplete, for higher-order programs.
